---
title: "Homework 3"
author: "Ying Jin"
date: "2020/10/6"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot_cpntinuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load the dataset

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user/ order variables -- user ID, order ID, order day and order hour. There are also item variables -- name, aisle, department, and some numeric codes.


How many aisles, and which are most products from?

```{r}
instacart %>% 
  count(aisle) %>%
  nrow()
  
instacart %>% 
  count(aisle) %>%
  arrange(desc(n))
```

There are 134 aisles and most products come from fresh vegetables.

Let's make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = forcats::fct_reorder(aisle,n)
  ) %>% 
  ggplot(aes(x = aisle, y =n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5,hjust = 1))
```

Three most popular items in three aisles

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples vs. Ice Cream

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>% 
  group_by(product_name,order_dow) %>% 
  summarise(
    mean_hour = mean(order_hour_of_day)
  ) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```

## Problem 2

First, load the accelerometer dataset and tidy it.

```{r load_tidy_accel, warning = FALSE, message= FALSE}
accel_df = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = starts_with("activity_"),
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "n_activ"
  ) %>% 
  mutate(
    weekday_vs_weekend = case_when(
      day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")~"weekday",
      day %in% c("Saturday","Sunday")~"weekend"
    ),
    day = factor(day),
    day = fct_relevel(day, c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
    minute = as.double(minute)
  )

```

The steps I took to wrangle the dataset include:

* Clean the column names;

* Change the dataset from wide format into long format;

* Create a new column named `weekday_vs_weekend` to indicate whether a certain day is a weekday or a weekend;

* Change the `day` variable into factor and relevel it;

* Change the `minute` variable into double.

The resulting dataset contains information about everyday "activity counts" in each commonly one-minute intervals of a 63 year-old male diagnosed with CHF. And the time range is five weeks. 

* It contains `r nrow(accel_df)` rows and `r ncol(accel_df)` variables.

* The dataset provides five-week observations on "activity counts" of a man. The counting starts at the midnight of each day and "activity counts" in each observations in each days are given. Also, information about which week a certain day belongs to, day of the week and whether that day is a weekday or weekend can be obtained.

Then I aggregate across minutes to create a total activity variable for each day and create a table.

```{r day_total_activ_tb}
accel_df %>% 
  group_by(week, day) %>% 
  summarise(
    total_activ = sum(n_activ)
      ) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activ
  ) %>% 
  knitr::kable()
```

According to the resulting table, this man is less active on Monday and active on the other days in the first two weeks. He is less active on weekends, especially on Saturday, in the last two weeks. 

Next I make a single-panel plot that shows the 24-hour activity time courses for each day. To better indicate the time of a day, I change the labels of x axis into hours.

```{r}
accel_df %>% 
  ggplot(aes(x = minute, y = n_activ, color = day,group = week)) +
  scale_x_continuous(
    breaks = c(240,480,720,960,1200,1440),
    labels = c("4","8","12","16","20","24")
  ) +
  labs(
    title = "Activity Counts for Each Day",
    x = "time (hour)",
    y = "activity counts" 
  ) +
  geom_point(size = 1) +
  geom_line()
```

Based on the plot above, it can be seen that in general, the man is most active at night (about 20 to 22 o'clock) and least active after 22 o'clock until the next day morning (about 6 o'clock). We can also see that he is very active in the noon, about 11 to 12 o'clock, on Sunday. Somtimes  the man is very active at other time, like about 7:00 on Thursday, 7:30 on Sunday, and 16:00 - 17:00 on Saturday and Sunday.

## Problem 3

Load NY NOAA dataset

```{r}
data("ny_noaa")
```
 
We can see that:

* This dataset provides information of total daily precipitation, snowfall, snow depth, maximum and minimum temperature for all New York state weather stations on each day from January 1, 1981 through December 31, 2010; 

* NY NOAA dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)
` variables;

* Key variables include weather station ID, date, snowfall, snow depth, maximum and  minimum temperature;

* There are NAs in records of weather stations,especially in `tmin` and `tmax`.

Then clean the dataset.

```{r}
ny_noaa_df = 
  ny_noaa %>% 
  mutate(
    tmin = as.numeric(tmin),
    tmax = as.numeric(tmax),
    tmax = tmax/10,
    tmin = tmin/10
  ) %>% 
  separate(date,into = c("year","month","day"),sep = "-") %>% 
  mutate(
    month = as.numeric(month),
    month = month.abb[month],
    ) 
```

The cleaning steps include:

* Convert `tmin` and `tmax` into number;

* Change the units of `tmax` and `tmin` into degree C;

* Create separate variables `year`, `month` and `day` from `date`;

* Change the values of `month` into corresponding month abbreviation.

Then get the most frequently observed value for snowfall.

```{r}
ny_noaa_df %>% 
  filter(!(is.na(snow))) %>% 
  group_by(snow) %>% 
  summarise(n_obs = n()) %>% 
  arrange(desc(n_obs))
```

It can be seen that the most common value of snowfall is 0 mm. That's because most of time there is no snowfall in NY state.

Next, I make plot of `year` vs average `tmax` oberved in differnt stations in January and July separately.

```{r}
ny_noaa_df %>% 
  filter(
    month %in% c("Jan","Jul"),
    !(is.na(tmax))
    ) %>% 
  group_by(id,year,month) %>% 
  summarise(
    mean_tmax = mean(tmax,na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id, color = id)) +
  geom_point() +
  geom_line() +
  facet_grid(month~.) +
  labs(
    title = "Average Maximum Temperature in Jan and Jul over Years",
    x = "year",
    y = "average maximum temperature(C)"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90, vjust = .5,hjust = 1)
    )
```

Below are interpretions of the plot:

* Apparently, the average maximum temperature in July is always higher than January;

* It can be noticed that the average maximum temperature in January is more variable than July;

* There are outliers of the average maximum temperature in both January and July. For example, in July, 1988, the average of maximum temperature recorded by one station is very low compared to others. And in January, 2005, there is also a very low average maximum temperature recorded by one station.

Then I make a plot showing `tmax` vs `tmin`.

```{r}
tmax_plot = 
  ny_noaa_df %>% 
  filter(
    !(is.na(tmax))
  ) %>% 
  ggplot(aes(y = tmax)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Maximum Temperature",
    y = "maximum temperature(C)"
  ) +
  scale_y_continuous(
    limits = c(-60,60)
  ) +
  guides(
    x = "none"
  )

tmin_plot = 
  ny_noaa_df %>% 
  filter(
    !(is.na(tmin))
  ) %>% 
  ggplot(aes(y = tmin)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Minimum Temperature",
    y = "minimum temperature(C)"
  ) +
  guides(
    x = "none"
  )

tmax_plot + tmin_plot
```

